---
title: "Data Wrangling I - Data Import"
author: Pavithra Srinivasan
date: September 19, 2023
output: github_document
---

```{r, setup}
library(tidyverse)
```

## Reading and Importing Data

Lets import the 'FAS_litters.csv' csv using a relative path. 
```{r}
litters_df = 
  read_csv("data/FAS_litters.csv")

## this will clean up the data and make it more readable

litters_df = 
  janitor::clean_names(litters_df)
  
```
## Now lets take a look at the clean data

Printing in the console. 

```{r}
litters_df
```


```{r}
## this shows the top and bottom few rows of the dataset. 
head(litters_df)
tail(litters_df)
```
```{r}
skimr::skim(litters_df)
```

## Other options to read_csv (looking at the help file)

```{r}
litters_df = read_csv("data/FAS_litters.csv", skip = 10, col_names = FALSE)
```

Lets look at NA values. 
```{r}
litters_df = read_csv("data/FAS_litters.csv", na = c("NA", 19))
```

Check out: '?read_csv()' for more information. 

## Learning Assessment
Repeat the data import process for pups. 

```{r}
pups_df = read_csv("data/FAS_pups.csv")
pups_df = janitor::clean_names(pups_df)
```

## Lets read other file formats. 

Read an excel file. 
```{r}
## first load the package (can also load it under the first code chunk)
library(readxl)

#then set up a relative path
mlb_df = read_excel("data/mlb11.xlsx")

# check if it worked
mlb_df
```
Read a SAS file. 
```{r}
library(haven)
pulse_df = read_sas("data/public_pulse_data.sas7bdat")
pulse_df
```

## Comparison with BASE R
read.csv instead of read_csv
important to note: Dont use read.csv. This is just for comparison. 

```{r}
litters_base = read.csv("data/FAS_litters.csv")
litters_readr = read_csv("data/FAs_litters.csv")

litters_base
litters_readr

## base r is just more difficult to look at, and it shows the whole data set without proper organization. 
```

## Exporting Data to a csv. 

Export the mlb sub-table. 

```{r}
write_csv(mlb_df, "data/mlb_subtable.csv")
```



